\Large \bf Relationship of Weather and Maize Yields in Kenya

### 2018-11-16 writing/draft1/draftMonika9.tex
working hardly
writing bullet points of the future paper..

\begin{itemize}
\item Seasonal cumulative rainfall
\item Seasonal standard deviation 
\item Seasonal coefficient of variation 
\item Maximum length of dry spell in number of days
\item ...\textcolor{blue}{the list to be finished}
\item...
\end{itemize}
\item Temperature
\begin{itemize}
\item
\item\textcolor{blue}{the list to be finished}..
\item
\end{itemize}
\end{itemize}
\subsection{Data}
\textcolor{blue}{source of data, resolution and plots/barplots of yield, precip and temp when final data available.}\\
\textcolor{red}{maybe a table of descriptive statistics or ref to the tale to appendix}
\textcolor{blue}{also how the weather data were aggregated spatially over counties}
\subsection{Linear mixed models}
\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}. As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}. Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset. These types of models include both fixed effects parameters and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}. On the other hand, random effect are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9, KERYch12}).

@@

\large Findings:
\begin{itemize}
\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important

\item Mar-Sept last year temperature very important for maize yields
\item SD temperature last year positive and significant
\item dry spell 20 MAM last year important (but not dry spell MAM10)
\item interesting. Precipitation 2 months MAM last year very significant and positive
\item mean temp last year negative and significant, hill shaped 
\end{itemize}

\underline{New findings:}

\begin{itemize}
\item The yields seem to be more responsive to weather on west than on east
\end{itemize}
\normalsize
\FloatBarrier
\pagebreak

@@

{\hspace{-1.5cm}
\begin{threeparttable}

\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(0,0) errors}}
% AmBEST00, AmBEST, AmBEST99

\label{Models1}

\begin{small}
\begin{tabular}{lcrcrcr}

\hline \vspace{-0.2cm} \\
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{Model 1}} &\multicolumn{2}{c}{\textit{Model 2}} &\multicolumn{2}{c}{\textit{Model 3}}\\
    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&Estimate&Sum sq.&Estimate&Sum sq.&Estimate&Sum sq.\\
 \hline 
\hline
\\
\vspace{-0.2cm}Intercept&$0.298$ $(0.066)^{***}$&---&$0.256$ $(0.061)^{***}$&---&$0.243$ $(0.076)^{**}$&---\\
  \\
\vspace{-0.2cm}Prec. total&$0.289$ $(0.045)^{***}$&$11.9$&$0.131$ $(0.046)^{**}$&$1.3$&$0.137$ $(0.125)^{}$&$0.3$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$-0.071$ $(0.028)^{*}$&$1.8$&$-0.075$ $(0.026)^{**}$&$1.5$&$-0.073$ $(0.039)^{\bullet}$&$1.0$\\
  \\
  \vspace{-0.2cm}Temp. - average&$-0.202$ $(0.062)^{**}$&$3.1$&$-0.188$ $(0.058)^{**}$&$2.9$&$-0.464$ $(0.144)^{**}$&$2.9$\\
  \\
    \vspace{-0.2cm}Prec. $\times$ temp.&---&---&$-0.135$ $(0.037)^{***}$&$2.5$&$-0.108$ $(0.051)^{*}$&$1.3$\\
  \\
      \vspace{-0.2cm}Prec. c. of variation&---&---&$-0.198$ $(0.035)^{***}$&$7.6$&$-0.181$ $(0.075)^{*}$&$1.7$\\
  \\
        \vspace{-0.2cm}Temp. std. dev.&---&---&$0.064$ $(0.021)^{**}$&$1.3$&$0.058$ $(0.027)^{*}$&$1.3$\\
  \\
          \vspace{-0.2cm}Prec. first 2 months&---&---&---&---&$-0.003$ $(0.040)^{}$&$0.001$\\
  \\
         \vspace{-0.2cm}Dry spell -length&---&---&---&---&$-0.050$ $(0.031)^{}$&$0.7$\\
  \\
        \vspace{-0.2cm}Dry spells $>9$ days&---&---&---&---&$0.018$ $(0.032)^{}$&$0.09$\\
  \\
          \vspace{-0.2cm}Max precipitation&---&---&---&---&$-0.008$ $(0.064)^{}$&$0.005$\\
  \\
          \vspace{-0.2cm}Degree days&---&---&---&---&$-0.016$ $(0.092)^{}$&$0.009$\\
  \\
          \vspace{-0.2cm}Heat wave d. ($>35$\degree C)&---&---&---&---&$0.032$ $(0.039)^{}$&$0.2$\\
  \\
          \vspace{-0.2cm}Max temperature&---&---&---&---&$0.278$ $(0.136)^{*}$&$1.2$\\
  \\
  \hline
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
\vspace{-0.2cm}
\\
\hline
\\
  \vspace{-0.2cm}Intercept\\
  \vspace{-0.1cm} \\ 
  \hline
    \vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{AIC:}}&$2114.98$&&$2144.88$&&$2118.15$\\

\hline
\vspace{-0.2cm}
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
        \begin{adjustwidth}{1cm}{} 
    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

{\hspace{-1.5cm}
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(2,1) errors}}
%AmArE2100, AmArE21
\label{ARMA21} 
\begin{footnotesize}

\begin{tabular}{lcrcrcr}

\hline \vspace{-0.2cm} \\
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{Model 1}} &\multicolumn{2}{c}{\textit{Model 2}} &\multicolumn{2}{c}{\textit{Model 3}}\\
    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&Estimate&F-value&Estimate&F-value&Estimate&F-value\\
 \hline 
\hline
\\
\vspace{-0.2cm}Intercept&$0.299$ $(0.064)^{***}$&$13.9$&$0.240$ $(0.064)^{***}$&$14.0$&$0.270$ $(0.074)^{***}$&$13.8$\\
  \\
\vspace{-0.2cm}Prec. total&$0.273$ $(0.042)^{***}$&$63.6$&$0.159$ $(0.051)^{**}$&$64.7$&$0.219$ $(0.117)^{\bullet}$&$65.2$\\
  \\
  \vspace{-0.2cm}Prec. total sq.&$-0.066$ $(0.026)^{*}$&$9.9$&$-0.079$ $(0.028)^{**}$&$10.1$&$-0.094$ $(0.037)^{*}$&$10.2$\\
  \\
  \vspace{-0.2cm}Temp. - average&$-0.177$ $(0.061)^{**}$&$8.4$&$-0.191$ $(0.062)^{**}$&$11.8$&$-0.380$ $(0.133)^{**}$&$8.7$\\
  \\
    \vspace{-0.2cm}Prec. $\times$ temp.&---&---&$-0.135$ $(0.037)^{**}$&$4.6$&$-0.101$ $(0.049)^{*}$&$7.5$\\
  \\
      \vspace{-0.2cm}Prec. c. of var.&---&---&$-0.139$ $(0.040)^{***}$&$11$&$-0.091$ $(0.070)^{}$&$11.1$\\
  \\
        \vspace{-0.2cm}Temp. std. dev.&---&---&$0.038$ $(0.024)^{}$&$2.4$&$0.038$ $(0.026)^{}$&$2.4$\\
  \\
          \vspace{-0.2cm}Prec. 2 months&---&---&---&---&$-0.009$ $(0.037)^{}$&$0.3$\\
  \\
         \vspace{-0.2cm}Dry spell -length&---&---&---&---&$-0.045$ $(0.030)^{}$&$2.2$\\
  \\
        \vspace{-0.2cm}Dry spells $>9$ days&---&---&---&---&$0.008$ $(0.031)^{}$&$0.1$\\
  \\
          \vspace{-0.2cm}Max precipitation&---&---&---&---&$-0.035$ $(0.060)^{}$&$0.2$\\
  \\
          \vspace{-0.2cm}Degree days&---&---&---&---&$-0.015$ $(0.088)^{}$&$0.6$\\
  \\
          \vspace{-0.2cm}Heat wave d.&---&---&---&---&$0.040$ $(0.037)^{}$&$3.1$\\
  \\
          \vspace{-0.2cm}Max temperature&---&---&---&---&$0.207$ $(0.123)^{\bullet}$&$2.8$\\
  \\
  \hline
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
\vspace{-0.2cm}
\\
\hline
\\
  \vspace{-0.2cm}Intercept\\
  \vspace{-0.1cm} \\ 
  \hline
  \vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{AIC:}}&$2064.54$&&$2064.99$&&$2099.90$\\

\hline
\vspace{-0.2cm}
\end{tabular} 
\end{footnotesize}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
        \begin{adjustwidth}{1cm}{} 
    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

>>>>

\begin{itemize}
\item Seasonal cumulative rainfall
\item Seasonal standard deviation 
\item Seasonal coefficient of variation 
\item Maximum length of dry spell in number of days
\item ...\textcolor{blue}{the list to be finished}
\item...
\end{itemize}
\item Temperature
\begin{itemize}
\item
\item\textcolor{blue}{the list to be finished}..
\item
\end{itemize}
\end{itemize}
\subsection{Data}
\begin{itemize}
\item \textcolor{blue}{Source of data}
\item \textcolor{blue} {Verbal description of the data, resolution, period, frequency}
\item \textcolor{blue} {Plot(s)/barplots of yield, precip and temp when final data available}
\item \textcolor{blue}{Maybe a table of descriptive statistics or ref to the tale to appendix}
\item \textcolor{blue}{Description how the weather data were aggregated spatially over counties}
\item \textcolor{red}{Description how the weather data were aggregated temporally - seasons}
\end{itemize}
\subsection{Linear mixed models}
\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}. As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}. Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset. These types of models include both fixed effects parameters and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}. On the other hand, random effect are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9, KERYch12}).

@@

\color{blue}
\begin{itemize}

\item Tables of estimates of:

\begin{enumerate}
\item Basic model which only includes seasonal precipitation and average seasonal temperature
\item The preferred specification which includes the significant weather aggregates (see the preliminary results in Table \ref{KenARe11})

\begin{itemize}
\item For all counties and also separately for Arid or semi-araid (ASAL) and non-ASAL
\end{itemize}

\end{enumerate}
\item Verbal description and interpretation of the results in table
\item Discussion and Conclusion:
	\begin{itemize}
	\item Stressing the importance of not only focusing on prediction of rain totals, but also other characteristics of the weather - particularly the variables found significant
	\end{itemize}
	\item \textcolor{red}{Caveats}
\end{itemize}
\color{black}

\FloatBarrier
\pagebreak

@@
